---
title: "data_analysis"
output: html_document
date: "2024-03-12"
---

# What can be said about the quality of the data?

To assess this, I propose to perform some sanity checks. First, we'd like to look at the distributions of all individual channels, and compare these between replicate. This will enable us to get a sense of what the phenotypes of these cells look like, in terms of central tendencies, and variability. Also, that will enable us to control for reproducibility.

# 1D Distributions

First, let's write a function that plots all desired distributions.
```{r,warning=FALSE}
plot_all_1D_distributions <- function(ini,end,condition){
  cyto_plot_save(sprintf("./figures/1d_distributions_%s.pdf",condition),width=25,height=25)
  cyto_plot_custom(layout=c(6,5))
  
  xlim_inf <- rep(c(0),times=9)
  xlim_sup <- c(4e6,4e6,1e6,1e6,1e6,7e5,2e6,7e5,1e6)
  
  for (i in 1:9){
    cyto_plot(fs[ini:end],channels=cyto_channels(fs)[i],density_stack=0.7,axes_trans=transList,title=condition,xlim=c(xlim_inf[i],xlim_sup[i])
              )}
  for (i in 10:30){
    cyto_plot(fs[ini:end],channels=cyto_channels(fs)[i],density_stack=0.7,axes_trans=transList,title=condition,xlim=c(-1e5,1e6))}
  cyto_plot_complete()}
```

Then we plot all 1D Distributions for all "postSpn","preSpn" and "saline" replicates.
```{r}
plot_all_1D_distributions(1,6,"postSpn")
plot_all_1D_distributions(7,12,"preSpn")
plot_all_1D_distributions(13,18,"saline")
```

Overall, postSpn samples are more reproducible than preSpn and saline. That also can be explained by the actual number of cells there are in each sample. There are 10 times more cells in postSpn versus preSpn, and 20 times more cells in postSpn versus saline.

```{r}
cyto_stats_compute(fs,stat="count")
```

Let's compare these distributions in each channel: pre|saline|post, to determine where distributions have changed.

```{r}
plot_1D_distribution_comparison <- function(){
  cyto_plot_save("./figures/1d_distribution_comparison.pdf",width=15,height=45*3)
  cyto_plot_custom(layout=c(30,3))
  
  xlim_inf <- rep(c(0),times=9)
  xlim_sup <- c(4e6,4e6,1e6,1e6,1e6,7e5,2e6,7e5,1e6)
  
  for (i in 1:9){
    cyto_plot(fs[7:12],channels=cyto_channels(fs)[i],density_stack=0.7,axes_trans=transList,title="preSpn",xlim=c(xlim_inf[i],xlim_sup[i]))
    cyto_plot(fs[13:18],channels=cyto_channels(fs)[i],density_stack=0.7,axes_trans=transList,title="saline",xlim=c(xlim_inf[i],xlim_sup[i]))
    cyto_plot(fs[1:6],channels=cyto_channels(fs)[i],density_stack=0.7,axes_trans=transList,title="postSpn",xlim=c(xlim_inf[i],xlim_sup[i]))
    }
  for (i in 10:30){
    cyto_plot(fs[7:12],channels=cyto_channels(fs)[i],density_stack=0.7,axes_trans=transList,
              xlim=c(-1e5,1e6),title="preSpn")
    cyto_plot(fs[13:18],channels=cyto_channels(fs)[i],density_stack=0.7,axes_trans=transList,
              xlim=c(-1e5,1e6),title="saline")
    cyto_plot(fs[1:6],channels=cyto_channels(fs)[i],density_stack=0.7,axes_trans=transList,
              xlim=c(-1e5,1e6),title="postSpn")}
  cyto_plot_complete()}

plot_1D_distribution_comparison()
```

# 2D distributions

Let's make one PDF per file.

```{r}
plot_2d_distributions <- function(k){
  xlim_inf <- rep(c(0),times=9)
  xlim_sup <- c(4e6,4e6,1e6,1e6,1e6,7e5,2e6,7e5,1e6)

  condition <- cyto_details(fs)$conditions[k]
  cyto_plot_save(sprintf("./figures/2d_heatmap_%s.pdf",condition),width=135,height=135)
  cyto_plot_custom(layout=c(30,30))
  for(i in 1:30){
    for(j in 1:30){
      cyto_plot(fs[k],channels=cyto_channels(fs)[c(i,j)],
                parent="root",
                xlim=c(lim_inf[i],lim_sup[i]),
                ylim=c(lim_inf[j],lim_sup[j]),
                title=cyto_details(fs)$conditions[k],
                axes_trans=transList)
    }
  }
  cyto_plot_complete()
}

for(p in 1:18){
  plot_2d_distributions(p)}
```

There is definitely a lot of things in there. Let's go line by line.
## postSpn79 
- Against all other scattering properties, nothing interesting to say. One population, that's it. Which also means that in terms of granularity and size, all these cells make a continuum. Then the variance it self, is hard to judge for me.
- Going along *FSC-A*:
  - *B220*: a B220 negative and a B220 positive populations. The positive population is supposed to be B2 B Cells. Naive and resident memory B cells. There indeed seems to be an enrichment for these in the sample compared to controls. But we can also find these cells in the control.
  - *CD8/IgD*: CD8/IgD neg and pos populations but no difference compared to controls.
  - *CD3*: small tail as well, but no big differences when looking at the 1d histograms, that are characterized by high variability.
  - *CD4*: clear bimodal distribution (as in saline and pre). But large part of the population is negative. Could there be an over compensation issue?
  - *GL7*: Tail towards GL7+ populations. True for saline and pre...
  

- Going vertical along *PD1*: something interesting. Distribution looks slightly bimodal. Turns out, population can be exploded:
    - B220- is slightly higher in PD1, than B220+.
    - CD4+ is slightly higher in PD1 than CD4-
    - Possibly not enough correction from PD1 to CD44 channels (BV510-A to BV570)
    - CD19- cells are slightly higher in PD1 than CD19+ cells.
    - Possibly not enough correction from PD1 to CD43 channels (BV510-A to BV750)
    - CD38- cells slightly lower than CD38+ cells in PD1.
    
- Going vertical along *CD44*: -/+ populations
    - B220: CD44+B220- and CD44-B220+ which makes sense as CD44 is supposed to be a marker of memory T-cells.
    - CD8_IdG: same, but number of CD8+ positive cells is very small.
    - CD3: CD44+ are also CD3+. But this is not obvious on the final phenograph (see cluster 20 and 12).
    - CD4: CD44+ are also CD4+. But there are counter example on the final phenograph (see cluster 10, naive T-cells, but how many are there?).
    - Possibly not enough correction from PD1 to CD44 channels (BV510-A to BV570)?
    - CD19- are CD44+, CD19+ are CD44-. Coherent with Phenograph.
    - CD62L+ cells are CDD44- (very few cells)
    - CD43: Positive correlation (same cluster on the phenograph)
    - CD11a: Positive correlation (same cluster on the phenograph)
    - *CD69*: interesting and somewhat complex: lots of CD69-CD44- cells, CD69+CD44+ cells (Trm cells), and also some CD44-CD69+ (Brm cells and naive/activated B cells)
    - *CD73*: same as CD69, similar pattern: lots of CD44+CD73+, CD44-CD73-, and some CD44-CD73+. But never off together.
    - *CD38*: CD44-CD38+, CD44+CD38-. But no CD38 on the phenograph.
    - IgM: IgM+ cells are CD44-. But there is cluster 12 (a few cells that are on the top of the tail?)

- Going vertical along *CD19*:
  - B220: B220+GL19+ cells (CD19+ are actually B cells, this is the criteria used to filtering T cells). But... there are some CD19+ cells, that are not B220+.
  - 2 clusters: 4 and 12. The main difference between these two? CD44 expression. 4 is supposed to be B1 Brm cells (but, the reason escapes me), whereas 12 is supposed to be B1 Bmem cells. What's the difference?
  - CD8_IgD: 4 populations there: CD19-CD8+ (T8 Trm cells), CD19-CD8-(Various Tcells), CD19+CD8_IgD+(which are also IgM+, Bcells naive or not), CD19+CD8_IgD-(activated IgM+ Bcells).
  - CD3+ cells are CD19-. Only Tcells.
  - CD4+ cells are CD19 (vast majority)
  - GL7: GL7 cells can be either CD19+ or CD19-
  - CD62L: CD62L cells can be either CD19+ or CD19-
  - CD43: very few cells are CD43+CD19+. Most of CD43+ are CD19- (marker of T-cells activation).
  - CD11: strong correlation with CD19+ cells. CD19- cells (T-cells) are almost always CD11+.
  - CD69: 4 populations
    - CD19+CD69-: B cells
    - CD19+CD69+: Brm cells
    - CD19-CD69-: T cells (non Trm)
    - CD19+CD69+: Trm cells
  - CD73: Most of the cells are CD19+CD73- or CD19-CD73+ (as visible in the phenograph). But there are cells in the 3 quadrants.
  - CD38: all CD19+ are CD38+. Continuum CD19-CD38- to CD38+.
  - IgM: diverse degree of IgM+ for CD19+ (Bcells)
  
- Going vertical along *CD62L*. Seems interesting. It is only on for a cluster of CD4+ T Naive Cells (cluster 10).
  - CD4: CD62L+ are supposed to be CD4+ cells (according to phenograph). It does not show when plotting CD62L versus CD4. It seems that most of the cells that are CD62L+, are also CD4- (and CD3-, and CD44-). It does not show, because most of these cells are scattered in many different cluster I feel.
  - CD19:  4 populations. CD19-CD62L+ (cluster 10), CD19+CD62L+ (spread in a bunch of Bcells cluster).
  - 
  
  
  
    
  
    


    
    
    
    
    
    
    





# 2D distributions

```{r}
names_list <- simplify(new_colnames)

toplot <- mydata %>% 
  dplyr::filter(condition=="postSpn") %>% 
  select(names_list[c(1:30)],"replica_old") %>% 
  pivot_longer(cols=names_list[c(1:30)],names_to = "var",values_to = "val") %>% 
  nest_by(var) %>% {
    d <- tidyr::expand(.,V1=var,V2=var)
    d <- left_join(d,mutate(.,data=list(rename_with(data,.fn=~paste0(.x,"_x")))), by = c("V1"="var"))
    d <- left_join(d,mutate(.,data=list(rename_with(data,.fn=~paste0(.x,"_y")))), by = c("V2"="var"))
    d
  } %>% 
  unnest(c(data.x,data.y))
  
ggplot(toplot, aes(x = val_x, y = val_y, color = replica_old_x)) +
  geom_point() +
  facet_grid(rows = vars(V1), cols = vars(V2))
```

Now let's systematically compare variables 2 by 2. Gates are already set, we need to merge.

```{r}
merged_fs <- cyto_merge_by(fs,"treatment")
```

```{r,ouput=}
lim_inf <- c(rep(c(0),times=9),rep(-1e5,times=20))
lim_sup <- c(4e6,4e6,1e6,1e6,1e6,7e5,2e6,7e5,1e6,rep(1e6,times=20))
replica <- cyto_details(fs)$replica
for(i in 1:18){
  cyto_gate_draw(fs[[i]],alias=sprintf("replica_%s",replica[i]),channels=cyto_channels(fs)[c(1,4)],xlim=c(lim_inf[1],lim_sup[1]),ylim=c(lim_inf[4],lim_sup[4]),type="boundary")
  Sys.sleep(5)}
```

```{r}
cyto_gate_draw(fs[[1]],parent="root",alias=sprintf("replica_%s",replica[i]),channels=cyto_channels(fs)[c(1,4)],xlim=c(lim_inf[1],lim_sup[1]),ylim=c(lim_inf[4],lim_sup[4]),type="boundary",popup=TRUE)
```

```{r}
cyto_plot_save("./figures/2d_distributions_postSpn.pdf",width=5,height=5)
cyto_plot_custom(layout=c(1,2))
lim_inf <- c(rep(c(0),times=9),rep(-1e5,times=20))
lim_sup <- c(4e6,4e6,1e6,1e6,1e6,7e5,2e6,7e5,1e6,rep(1e6,times=20))
#How to fuse populations in a single one, while gating them.

for(i in 1:1){
  for(j in 1:2){
    cyto_plot(merged_fs[[1]],channels=cyto_channels(fs)[c(i,j)],xlim=c(lim_inf[i],lim_sup[i]),ylim=c(lim_inf[j],lim_sup[j]),point_col)
  }
}
cyto_plot_complete()

```



## Scattering distributions
Let's focus on the scatter data first.
We first prepare the data to be able to use facet_wrap.
```{r}
measurement <- c('FSC.A','FSC.H','FSC.W','SSC.A','SSC.H','SSC.W')
variable <- c('replica','replica_old',"condition")
toplot <- mydata %>% select(all_of(c(measurement,variable)))
toplot <- pivot_longer(toplot,cols = measurement,names_to = 'scattering',values_to = 'values')
```

```{r}
toplot %>% ggplot()+
  geom_freqpoly(aes(values,after_stat(density),col=replica),bins=100)+
  facet_wrap(condition~scattering,scales="free")+
  theme_cowplot()+
  expand_limits(x = 0, y = 0)
```

```{r}
measurement <- c('FSC.A','FSC.H','FSC.W','SSC.A','SSC.H','SSC.W')
variable <- c('replica','replica_old',"condition")
toplot <- mydata %>% select(all_of(c(measurement,variable)))
#toplot <- pivot_longer(toplot,cols = measurement,names_to = 'scattering',values_to = 'values')
```

```{r}
toplot %>% ggplot()+
  geom_point(aes(FSC.A,SSC.A),alpha=0.1) + 
  facet_wrap(condition~replica,scales="free")+
  theme_cowplot()+
  expand_limits(x = 0, y = 0)

```

```{r}
toplot %>% ggplot()+
  geom_point(aes(FSC.A,FSC.H),alpha=0.1) + 
  facet_wrap(condition~replica,scales="free")+
  theme_cowplot()+
  expand_limits(x = 0, y = 0)
```


```{r}
toplot %>%
  filter(condition=="postSpn") %>% 
  ggplot()+
  stat_bin_2d(aes(FSC.A,FSC.H),bins=500) + 
  #facet_wrap(condition~replica,scales="free")+
  theme_cowplot()+
  expand_limits(x = c(0,5e6), y = c(0,4e6))
```

```{r}
mydata %>%
  filter(condition=="postSpn",replica_old==84) %>% 
  ggplot()+
  stat_bin_2d(aes(CD45.2.BUV737,CD45.AF532),bins=50) + 
  #facet_wrap(condition~replica,scales="free")+
  theme_cowplot()
  #expand_limits(x = c(0,5e6), y = c(0,4e6))
```

```{r}
mydata %>%
  filter(condition=="postSpn") %>% 
  ggplot()+
  stat_bin_2d(aes(CD45.2.BUV737,SevenAAD),bins=500) + 
  #facet_wrap(condition~replica,scales="free")+
  theme_cowplot()
  #expand_limits(x = c(0,5e6), y = c(0,4e6))
```

# Generate NxN plots of all 2D heatmaps

We want to plot all variables against each other using heat map fashion.
One PDF for each replica experiment. It will enable us to judge precisely which population was actually taken in the data, and which one was discarded.

Plus, we'll check potentially bad unmixing/weird populations.

To do that quick, the best is certainly to use ggplot. But there must be a way to do that quickly taking advantage of relevant Bioconductor packages.



```{r}




```

