---
title: "data_analysis"
output: html_document
date: "2024-03-12"
---

## Descriptive analysis
Let's first check how many cells there are per condition and replicate in these data.

```{r, }
summary <- mydata %>% group_by(condition,replica) %>% 
  dplyr::count()

print.data.frame(summary)
```

## Scattering distributions
Let's focus on the scatter data first.
We first prepare the data to be able to use facet_wrap.
```{r}
measurement <- c('FSC.A','FSC.H','FSC.W','SSC.A','SSC.H','SSC.W')
variable <- c('replica','replica_old',"condition")
toplot <- mydata %>% select(all_of(c(measurement,variable)))
toplot <- pivot_longer(toplot,cols = measurement,names_to = 'scattering',values_to = 'values')
```

```{r}
toplot %>% ggplot()+
  geom_freqpoly(aes(values,after_stat(density),col=replica),bins=100)+
  facet_wrap(condition~scattering,scales="free")+
  theme_cowplot()+
  expand_limits(x = 0, y = 0)
```

```{r}
measurement <- c('FSC.A','FSC.H','FSC.W','SSC.A','SSC.H','SSC.W')
variable <- c('replica','replica_old',"condition")
toplot <- mydata %>% select(all_of(c(measurement,variable)))
#toplot <- pivot_longer(toplot,cols = measurement,names_to = 'scattering',values_to = 'values')
```

```{r}
toplot %>% ggplot()+
  geom_point(aes(FSC.A,SSC.A),alpha=0.1) + 
  facet_wrap(condition~replica,scales="free")+
  theme_cowplot()+
  expand_limits(x = 0, y = 0)

```

```{r}
toplot %>% ggplot()+
  geom_point(aes(FSC.A,FSC.H),alpha=0.1) + 
  facet_wrap(condition~replica,scales="free")+
  theme_cowplot()+
  expand_limits(x = 0, y = 0)
```


```{r}
toplot %>%
  filter(condition=="postSpn") %>% 
  ggplot()+
  stat_bin_2d(aes(FSC.A,FSC.H),bins=500) + 
  #facet_wrap(condition~replica,scales="free")+
  theme_cowplot()+
  expand_limits(x = c(0,5e6), y = c(0,4e6))
```

```{r}
mydata %>%
  filter(condition=="postSpn",replica_old==84) %>% 
  ggplot()+
  stat_bin_2d(aes(CD45.2.BUV737,CD45.AF532),bins=50) + 
  #facet_wrap(condition~replica,scales="free")+
  theme_cowplot()
  #expand_limits(x = c(0,5e6), y = c(0,4e6))
```

```{r}
mydata %>%
  filter(condition=="postSpn") %>% 
  ggplot()+
  stat_bin_2d(aes(CD45.2.BUV737,SevenAAD),bins=500) + 
  #facet_wrap(condition~replica,scales="free")+
  theme_cowplot()
  #expand_limits(x = c(0,5e6), y = c(0,4e6))
```

# Generate NxN plots of all 2D heatmaps

We want to plot all variables against each other using heat map fashion.
One PDF for each replica experiment. It will enable us to judge precisely which population was actually taken in the data, and which one was discarded.

Plus, we'll check potentially bad unmixing/weird populations.

To do that quick, the best is certainly to use ggplot. But there must be a way to do that quickly taking advantage of relevant Bioconductor packages.



```{r}




```

